Predicting the effects of character displacement on food-web dynamics

### 2019-02-09 manuscript_ECD_model.Rmd
updates to in-text citations

Although explicit models of resources gave insights to the evolution of character displacement, the ecological feedback onto consumer-resource dynamics have received surprisingly little attention. 
This is likely because the ecological feedback has been primarily studied through the lens of competition and coexistence theory [@Lawlor1976; @Germain2018; @Bassar; @McPeek2019]. 
For example, early theoretical work by @Lawlor1976 showed that ecological character displacement promotes coexistence by favoring specialized consumers that experience reduced interspecific competition. 
<!--
On the other hand, specialization often increases a consumer’s ability to capture a specific resource.
From food-web theory, we know that increased feeding rates often lead to resource suppression, which if sufficient enough, can generate oscillations, and hence less stable, consumer-resource dynamics [Rosenzweig 1971; Murdoch et al. 2003; McCann 2011].
-->
Yet, this reduction in interspecific competition may, at the same time, increase the interspecific interaction between a consumer and the resource it specializes on.  
Both food-web theory and empirical studies have shown that increasing the strength of consumer-resource interactions often leads to resource suppression, which if sufficient enough, can generate oscillations, and hence less stable consumer-resource dynamics [@Rosenzweig1971; @Luckinbill1973; @Murdoch2002; @Murdoch2003; @McCann2011].
Thus, a food-web perspective, which accounts for both the direct and indirect effects of consumer-resource interactions, may yield new insight to the ecological consequences of character displacement.

@@

Here, I address this knowledge gap by building a mathematical model to examine how ecological character displacement affects consumer-resource dynamics in a food-web context. I address two questions: (1) How does ecological character displacement affect resource abundances? (2) How does character displacement affect food-web stability? To test the generality of these ecological effects, I examined three different foraging scenarios: *CONDENSE THIS. NO NEED TO GIVE ALL THE DETAILS YET. RIGHT NOW IT IS JUST CONFUSING.* (i) resources occur in the same habitat and consumers exhibit a linear functional response [e.g. @MacArthur1972; @Lawlor1976]; (ii) resources occur in different habitats and consumers exhibit a linear functional response [e.g. @Lawlor1976]; (iii) resources occur in distinct habitats and consumers exhibit a more realistic functional response [@McCann2005]. I also explore the contingency of these effects to different assumptions about evolutionary trade-offs in consumer foraging traits. I focus on the ecological side of these dynamics because the evolutionary aspects have been well characterized [@Lawlor1976; @Taper1985; @Abrams1986; McPeekRECENT]. I show that the consequences of character displacement on food-web dynamics are not always intuitive and depend on the foraging scenario and tradeoff in consumer traits.

>>>>

Although explicit models of resources gave insights to the evolution of character displacement, the ecological feedback onto consumer-resource dynamics have received surprisingly little attention. 
This is likely because the ecological feedback has been primarily studied through the lens of competition and coexistence theory [@Lawlor1976; @Germain2018; @Bassar2017; @McPeek2019]. 
For example, early theoretical work by @Lawlor1976 showed that ecological character displacement promotes coexistence by favoring specialized consumers that experience reduced interspecific competition. 
<!--
On the other hand, specialization often increases a consumer’s ability to capture a specific resource.
From food-web theory, we know that increased feeding rates often lead to resource suppression, which if sufficient enough, can generate oscillations, and hence less stable, consumer-resource dynamics [Rosenzweig 1971; Murdoch et al. 2003; McCann 2011].
-->
Yet, this reduction in interspecific competition may, at the same time, increase the interspecific interaction between a consumer and the resource it specializes on.  
Both food-web theory and empirical studies have shown that increasing the strength of consumer-resource interactions often leads to resource suppression, which if sufficient enough, can generate oscillations, and hence less stable consumer-resource dynamics [@Rosenzweig1971; @Luckinbill1973; @Murdoch2002; @Murdoch2003; @McCann2011].
Thus, a food-web perspective, which accounts for both the direct and indirect effects of consumer-resource interactions, may yield new insight to the ecological consequences of character displacement.

@@

Here, I address this knowledge gap by building a mathematical model to examine how ecological character displacement affects consumer-resource dynamics in a food-web context. I address two questions: (1) How does ecological character displacement affect resource abundances? (2) How does character displacement affect food-web stability? To test the generality of these ecological effects, I examined three different foraging scenarios: *CONDENSE THIS. NO NEED TO GIVE ALL THE DETAILS YET. RIGHT NOW IT IS JUST CONFUSING.* (i) resources occur in the same habitat and consumers exhibit a linear functional response [e.g. @MacArthur1972; @Lawlor1976]; (ii) resources occur in different habitats and consumers exhibit a linear functional response [e.g. @Lawlor1976]; (iii) resources occur in distinct habitats and consumers exhibit a more realistic functional response [@McCann2005]. I also explore the contingency of these effects to different assumptions about evolutionary trade-offs in consumer foraging traits. I focus on the ecological side of these dynamics because the evolutionary aspects have been well characterized [@Lawlor1976; @Taper1985; @Abrams1986; @McPeek2019]. I show that the consequences of character displacement on food-web dynamics are not always intuitive and depend on the foraging scenario and tradeoff in consumer traits.
